A Hitchhiker's Guide to Reproducible Research" # replace with something serious

### 2021-05-25 manuscript.Rmd
beginn rewrite, add introduction

# Why you should care about reproducibility

A cornerstone of scientific progress is the possibility to critically assess the validity of peer scientists’ scientific conclusions. <!-- CJ More fundamental: Reproducibility is a cornerstone of the scientific METHOD. I think it's more important to emphasize this, rather than critical assessment of others' work. See Karl Popper, The Logic of Scientific Discovery, Routledge, London, 1992, p. 66. --> 
Scientific claims should not be credible because of their originators' authority but by the transparency and replicability of their supporting evidence [@aac4716]. <!-- CJ This is a red herring; nobody said that an argument to authority should be the foundation of scientific evidence. I would suggest focusing on replicability as an essential component of the scientific method. -->
Here, *replicability* refers to the degree to which original studies and their derivatives yield consistent results.
In other words, a successful replication study finds a consistent effect with an original study in a new sample of data.
Enabling the assessment of scientific conclusions requires a systematic and rigorous approach to research.
Researchers across various fields have identified a *crisis* of non-replicability of findings [e.g. @aac4716]. <!--citation needed, e.g. [@aac4716]-->

<!-- define replicability and then... 
 define reproducibility -->
Even worse, researchers might often not even be able to achieve reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code. <!-- CJ I would introduce the topic of reproducibility differently. "Even worse" is editorializing; not appropriate for a scientific paper. Instead, consider saying: More fundamental even than replication across datasets is the reproducibility of findings within a single dataset. --> 
When analyses are not properly documented, small mistakes, or circumstances outside the researcher's control (e.g., a corrupted file, a coauthor leaving academia) may render them non-reproducible. 
Over the past decades, computational statistical methods have radically changed the ability of researchers from all disciplines to process and analyze data with complex statistical models. <!-- CJ I think it's more important to focus on the methods that have been developed to ensure reproducibility!)-->
With these advances came challenges that contribute to broader concerns of irreproducibility in the published literature, among them the lack of exact and unambiguous documentation of computational methods and each computational step taken.
With computationally intensive data analyses <!-- CJ The complexity of the analyses is orthogonal to the necessity to make them reproducible. -->, access to the steps taken to process data and generate findings is as important as access to the data themselves [@vanlissa2020worcs].
Only with access to both data and procedures <!-- CJ and software environment --> can we guarantee full computational reproducibility, meaning that the creations emanating from a given data analysis (a statistical result or an entire manuscript describing the results) can be exactly reproduced at any time in the future, by any person, and on any computer.

In fact, the research community has long accepted reproducibility as an important element of "good scientific practice." <!-- CJ: This is the crucial point. I would start the paper with this. See my reference to Popper in the opening sentence. -->
However, ensuring reproducibility of computational analyses is still tedious and difficult <!-- CJ I would not editorialize this. Moreover, this sounds discouraging. I would focus on what you have done to make reproducibility more attainable. -->, and there are still no widely accepted standards that can guide researchers when setting up their data analyses. <!-- CJ These are two separate points, each important enough to have their own sentence, with some further clarification. --> 
Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories <!-- CJ you have not introduced any open repositories yet. --> do not provide sufficient information to reproduce relevant computational and statistical results.
For example, @hardwicke2018 attempted to replicate open materials in the journal Cognition.
Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced —-- even with the original authors' assistance.
@obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).
Thus, there is a dire need to establish standards for reproducible workflows for scientific rigor.

From our personal experience, the lack of a proper incentive structure and the dreary perspective of having to spend several extra hours if not days to properly set of a code repository to become reproducible, makes many scholars skip this extra step.
If reproduction would require merely minutes instead of hours, it could greatly facilitate collaboration within and across research projects.
<!-- Guiding principle: spending machine compute time instead of human research time -->
To achieve this boost in productivity, we must place the burden of reproducibility upon computers instead of the researchers.
To this end, we demand that a scientific product (no matter whether this is a preliminary report or a full research article) should both be understandable by researchers and automatically reproducible by computers with minimal constraints to the researchers' own workflows.

<!-- Which problems are solved by which tool (like in Andreas talk)-->
<!--stolen from: https://brandmaier.github.io/reproducible-data-analysis-materials/KULeuvenQuantPsy2020.html#11 -->
We identified the following common four causes for non-reproducibility:

1.	Multiple inconsistent versions of code and data; for example, the dataset has changed over time because outliers were removed at a later stage, or the code has changed because a bug was removed at some point in time;
2.	Copy-and-paste errors; for example, results are manually copied from a computing environment into a text processor, causing inconsistencies between the reported result and the reproduced result;
3.	Ambiguous order of code execution; for example, with multiple data and code files, it may be unclear which scripts should be executed in what order;
4.	Broken dependencies; for example, a given analysis depends on a software package that might not be available on a different computer or no longer exists.

We have developed a workflow that leverages established tools and practices from software engineering to achieve long-term computational reproducibility of scientific data analyses resting on four pillars that address the aforementioned causes of non-reproducibility [@Peikert2019]:

1. Version control
2. Dynamic document creation
3. Dependency tracking
4. Software management

This tutorial follows the recommendations of @Peikert2019 and uses _Git_ for version control, _R Markdown_ for dynamic document creation, _Make_ for dependency tracking, and _Docker_ for software management.
The interplay of these four components is shown in Figure \@ref(fig:schematic).
However, which tools you use to implement these concepts is a matter of taste and project requirements, and other choices are possible depending on the context and the degree of reproducibility one aims for [e.g., @vanlissa2020worcs; @turingway2021turing].
At several points, we suggest other viable alternatives.

```{r schematic, eval = TRUE, echo = FALSE, fig.cap="Schematic illustration of the interplay of the four components (in dashed boxes) central to the reproducible workflow: version control (Git), dependency tracking (Make), software management (Docker), and dynamic document generation (R Markdown). Git tracks changes to the project over time. Make manages dependencies among the files. Docker provides a container in which the final report is built using dynamic document generation in R Markdown. Reproduced from @Peikert2019.  "}
# file gets downloaded in Makefile
knitr::include_graphics("images/nutshell.svg", auto_pdf = TRUE)
```

This tutorial guides the reader through each step of an open research project with an emphasis on guaranteeing reproducibility from the very first step on.
This paper was designed to help researchers to practically apply the TOP guidelines [except for number 8. which does not apply to single research projects, @nosekPromotingOpenResearch2015; @vanlissa2020worcs].
We start by sketching an initial idea in a repository shared with colleagues that potentially collaborate only remotely.
We continue with the creation of a concrete data analysis plan, which ideally is already implemented in code, and a dynamic report document that can be preregistered.
Once the empirical data has been collected, the results can be calculated and appear directly in the dynamically generated document, where ideally only the discussion section needs to be added before the whole archive can be permanently stored as a fully reproducible digital artifact using a unique DOI in a persistent location.

We have structured the tutorial with a _learning by doing_ approach in mind, such that readers can follow along at their own computers.
We explicitly encourage readers to try out all R commands for themselves.
Unless stated otherwise, all code blocks are meant to be run in the statistical programming language R (tested with version `r with(R.version, paste0(major, ".", minor))`).
To guide the reader through all steps of a research project, we will use a exemplary research question:

> "Is there a mean difference in the personality trait 'Machiavellism' between genders?"

This example serves only didactic purposes and not to derive substantive claims.
Nevertheless, everything else is kept as close to a real research question as possible and we have preregistered the analysis without looking at the data beforehand and use actual data subsequently.

Here, we present a workflow that has proven useful for our way of working with a strong focus on Open Science.
However, not all steps are necessary to achieve reproducibility.
We have marked those steps that can be omitted as "optional" in the section headings.

Table \@ref(tab:overview) shows all steps required to to create and open and reproducible research project as they are presented in the remainder of the manuscript.

```{r overview, echo=FALSE}
overview <-
  read_csv(
    here::here("data", "overview.csv"),
    col_types = "cc",
    col_names = c("Phase", "Steps"),
    trim_ws = FALSE
  )
pander::pander(
  overview,
  split.tables = Inf,
  split.cells = Inf,
  missing = "",
  keep.line.breaks = TRUE,
  style = "grid",
  justify = 'left',
  caption = "\\label{tab:overview} Overview over all steps to create and open and reproducible research project."
)
```

# Setup

<!-- keep it brief, let `repro` do the work -->

This tutorial is targeted at scholars with at least intermediate R knowledge and we assume that you have already installed R and RStudio (if not, check the `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/install.md", "installation guide")`).
To create reproducible research projects in R, we have implemented our workflow in the  [`repro`-package](https://github.com/aaronpeikert/repro)[^repropackage]:

[^repropackage]: https://github.com/aaronpeikert/repro

```r
# -- type this on the R console --
if(!requireNamespace("remotes"))install.packages("remotes")
remotes::install_github("aaronpeikert/repro")
library("repro")
```

The [`repro`-package](https://github.com/aaronpeikert/repro) contains several helpers to work with the suggested tools for reproducibility.
For example, the package can check if your computer has all the required software packages installed and ready.

```{r}
# -- type this on the R console --
check_git()
check_make()
check_docker()
```

```{r, eval=FALSE}
check_github()
```

If any of these commands find that a given software requirement is not installed or set up properly, you will get a message with detailed instructions (tailored to your operating system) on how to set up your computer.
Please make sure that all three commands pass before you continue with the tutorial.
This process is only required once per computer and does not need to be repeated for each project.

# Planning Phase

<!-- researcher should begin early to enjoy most benefits -->
Reproducibility is invaluable for effective collaboration---including with your "past self".
While you may be tempted to consider reproducibility only when you are close to completing a project, we believe that upfront investment pays off in your increased productivity.
Therefore, we suggest starting early, ideally from the moment you decide to pursue a project to incorporate tools to help you achieve reproducibility gradually.
The first step towards reproducibility is to bundle all required files in one folder.
This way, you never lose anything, and you can easily share the files without missing anything.
<!-- Hands-on: Introducing RProjects-->
RStudio's _R Projects_ facilitate this concept of a project folder.
We recomend the [`here`](https://github.com/r-lib/here)[^here] package [@R-here] to refer to files, because it works consistently across different operating systems, and between R scripts and [R Markdown].
The `repro` package offers an R Studio template, which sets up a file structure that follows best practices and conventions for placing scripts and data [@rextensions].
To create one, click on:

[^here]: https://github.com/r-lib/here

> File → New Project... → New Directory → New Project

## Markdown

<!--Keep notes.-->
 A great tool to capture ideas or meeting notes and eventually complete entire journal articles is Markdown.
 Markdown is a lightweight text format that is very close to plain text with a minimal subset of reserved symbols for formatting instructions.
 This way, Markup does not need any specialized software for editing, it is highly fool-proof (unlike for example LaTeX) and can be ported to various document formats, such as HTML Websites, a Microsoft Word document, a typeset PDF file (for example, via LaTeX journal templates), or a Powerpoint presentation.
We suggest to use Markdown for everything, starting from simple sketches of your ideas to your scientific manuscripts[@R-rticles] and presentations[@revealjs]---and even your CV [@vitae].
To create a Markdown document, click on:

> File → New File → Markdown File

Markdown is simply text, where some characters let you add a minimal amount of structure:

```{r, results='asis', echo=FALSE}
# in case we want to reuse this example later
markdown_example <- "
<!--this is a Markdown file -->
# Header

Normal text.
Important **word**.

To do list:

* do research
* do more research
* spend time with family
"
cat("```markdown", markdown_example, "```", sep = "")
```

We will later see how this simplicity enables us to create many different document formats (i.e. Word, PDF, HTML) and incorporate code (i.e. R, Python, Julia).

## Git

Another advantage of Markdown is that because of its likeness to plain text, you can easily version-control it with Git and leverage Gits features for tracking, displaying, and merging sequential and concurrent changes.
You may be familiar with Microsoft Words' "Track Changes"-Feature.
Version control is similar in its basic idea but more potent in its features.
Instead of tracking only one file, you can keep track of the whole project directory, and instead of losing all changes when you agree to them, the entire history is preserved transparently.
For example, this paper has been going through `r system2("git", c("rev-list",  "--count",  "HEAD"), stdout = TRUE)` `r link("https://github.com/aaronpeikert/repro-tutorial/commits/main", "iterations")`.
To activate Git in a given project, you can call:
<!-- Hands-on: Introduce Git -->

```r
# `package::function()` → use function from package without `library(package)`
usethis::use_git()
```

Recording each change you made helps you to iterate more quickly because you know that you can effortlessly go back to previous versions.
For example, you can write down a rough project idea, knowing your collaborators and you will iterate and improve.
<!-- This is maybe the place to introduce the example-->
First, you could create a new Markdown file named `idea.md` that looks like this:

```markdown
<!--this is a Markdown file -->

# Hypothesis

Machiavellism is higher in male persons.

# Analytic Strategy

Independent t-test

# Sample

Can we use openpsychometrics.org data?
```

And you could commit it in Git.
For convenience, we use the [`gert`-package](https://github.com/r-lib/gert) [@R-gert], which wraps most git commands into R functions.
Alternatively, we suggest to either work directly in the terminal (for example, RStudio offers a command shell with various options accessible in the Menu Tools → Terminal), or work using the Git pane in RStudio that offers buttons for basic operations.
<!--I am leaning towards doing everything in R, not sure though-->
<!--For Git we maybe should use RStudio's Git pane?-->

To commit your first changes to the project, start with staging all files to be included in the commit using `git_add()`.
Here, we stage our idea collection in the file `idea.md` and then commit it to our local repository with a meaningful commit message that will later help us keep track of all the changes over time.

```r
gert::git_add("idea.md")
gert::git_commit("add a first concept")
```

Now assume that after discussing with your colleagues, you settle onto a final analysis strategy using latent variable models with tests for measurement invariance because power is greater and you want to exclude bias in measurement as possible confounder [@flake2020].
This results in changes to your notes as follows:

```markdown
<!--this is a Markdown file -->

# Hypothesis

Machiavellism is higher in male persons.

# Analytic Strategy

Multigroup Confirmatory Factor Analysis with equal loadings and intercepts

# Sample

Can we use openpsychometrics.org data? 
```

Again, stage the changes to this file and commit like before using `git_add()` and `git_commit()` or use the following shorthand to stage all new and modified files and then commit:

```r
# -- type this on the R console --
gert::git_commit_all("switch to CFA to exclude bias in measurement as possible confounder")
```

And GitHub can show you what has changed, like in Figure \@ref(fig:idea-change).
This detection of what has changed works on a line-by-line basis.
We recommend therefore to keep each sentence in a Markdown on its own line to generate clean difference views.
GitHub is, in its essence, a web interface to Git, and you may learn more about how to use it under the section on [GitHub].

```{r idea-change-screenshot, include=FALSE}
# unfortunately css selectors won't work for pdf so we'll use png instead
idea_change <- here::here("images", "idea-change.png")
if(!file.exists(idea_change)){
  webshot2::webshot("https://github.com/aaronpeikert/repro-tutorial/commit/76f5d4e78b3d3bf22e84e9137756bc17767dff3b?diff=split",
                  file = idea_change,
                  selector = "div#diff-2f8beaa39e5c98706d8abb3361a8383776f8df8a195a21c0e6eeb25e5aa48f45", # only select file change
                  zoom = 2)# higher resolution
}
```

```{r idea-change, echo=FALSE, fig.cap="A screenshot of how a change tracked in Git is represented by GitHub.  ", out.width='100%'}
knitr::include_graphics(idea_change)
```

<!--I am troubled by how we should recommend to learn Git-->
Git can be challenging to learn, because it requires a rather complex mental model of maintaining potentially differing versions between different repositories (typically one central repository and one local repository per each author while each repository may have different branches).
However, there is a subsets of commands which we present here and in the section on [GitHub] (commit, push, pull) which enable you to use Git for simple version control.
There are many tutorials on learning Git and @vuorreCuratingResearchAssets2018 provide a good starting point.
However Git is designed for collaboration, we therefore recommend finding a partner to attend an online lecture or workshop to learn the ins and outs of Git.

## Simulation (optional)

<!-- impress with neatness → sample size planning, preregistration and analysis in one 😎-->
<!-- but explain that good science is often not as neat, strive for the ideal -->

```{r, include=FALSE}
source(here::here("R", "simulation_funs.R"))
```

```{r, include=FALSE}
loadings_jones_paulhus <- c(38, 31, 40, 52, 59, 71, 62, 46, 51)/100
cohend_jones_paulhus <- c(24, 29, 35)/100
```

Monte Carlo simulations are a crucial tool to investigate properties of a statistical model and they are routinely employed to develop or evaluate statistical models.
For example, @maccallum1999sample used Monte Carlo simulations to investigate optimal sample size for exploratory factor analysis, or @curran1996robustness studied the robustness of the $\chi^2$ test statistic under non-normal data. In Bayesian data analysis, prior predictive checks generate data from a model according to the prior distribution in order to asses whether a prior is appropriate [@gabry2019visualization].

While rather uncommon for an empirical research project, we argue that simulating data offers various benefits in the planning phase of a study.
First, you have to address and could plan for data analytic problems on a conceptual level before you made the considerable effort to acquire data [@axelrodAdvancingArtSimulation1997].
Second, if you only simulate a single dataset, you can perform a simple technical check, that is the analysis software outputs parameter estimates [@braiekTestingMachineLearning2020], e.g. when the simulated data contains missings.
Third, if you perform a Monte Carlo simulation, you may evaluate how the data analysis copes with expected data complications, e.g. non-normal data [@harrisonIntroductionMonteCarlo2010; @raychaudhuriIntroductionMonteCarlo2008].
Fourth, you can plan your resources efficiently, e.g. determine required sample size to achieve certain precision or power [@brandmaier2018precision].

A Monte Carlo simulation affords you an ideal planning opportunity.
We realize that this comes at an expensive but believe the reduction in the risk that the data is not sufficient to answer the research questions offsets the required effort.
This advantage is especially pronounced when you plan to preregister your study.
In fact, we argue that a Monte Carlo simulation can combine the analysis, sample size planning and preregistration (see [Preregistration as Code].

No matter which analytic strategy you plan, it will come with assumptions about the data generating process under investigation. 
A simulation forces us to deal with these assumptions and their consequences explicitly.
If you identify unrealistic assumptions, you can either deliberately violate these assumptions and, by simulation, check if the analytic strategy is robust or modify the modeling approach to explicitly address them.

Typically, you have to make a number of educated guesses to simulate data that closely matches your expectation.
These guesses strongly depend on the study you are planning, and there cannot be a one-fits-all solution.
Since we typically use parametric statistical models (general linear model or SEM) or hypothesis tests that can be expressed as Gaussian models, we can usually use these models directly to simulate data.
We will show an example using structural equation modeling (SEM) for our Machiavellism example.
Note that the general process applies to any statistical model, but the simulation here is specific to our hypothetical research question.

If you are unfamiliar with the specifics of SEM, you may briefly skim the next sections and continue in section ["Random Number Generation"](#prng).
To simulate data in SEM, one first specifies a model (typically in the form of a structural model and a measurement model), selects plausible population values for all parameters, computes the model-implied mean vector and covariance matrix, and then samples as many observations as needed from a multivariate normal distribution following with the given mean vector and covariance matrix.

The selected model parameters should express our expectations for the Machiavellism subscale of the short dark triad [SD3, @Jones2013].

As a starting point, we can use already published information.
@Jones2013 report a set of factor loadings from an Exploratory Structural Equation Model [@asparouhovExploratoryStructuralEquation2009].
The estimates for the construction sample, which was recruited from Amazon’s Mechanical Turk (MTurk) and consists of 489 adults from Canada and the United States, can be seen in Table \@ref(tab:loadings))
They report a gender difference (at the manifest level), which depending on the study ranges between `r min(cohend_jones_paulhus)` and `r max(cohend_jones_paulhus)`.

```{r loadings, echo=FALSE}
knitr::kable(t(data.frame(Item = as.character(1:9),
                        Loading = loadings_jones_paulhus)),
             caption = "Factor loadings from an Exploratory Structrural Equation Model as reported in Jones \\& Paulhus (2013)")
```

Typically, it is recommended to choose conservative values in the study planning phase [@brandmaier2018precision].
This way, we make sure that our analytic strategy is viable even in non-optimal circumstances, for example if the measurement instrument works less well then expected or the population characteristics are more different than expected.
As a heuristic guess, we assume that the factor loadings are 30% lower than those reported by @Jones2013 and that the standardized mean difference is only 0.2.
These assumption can be simply translated into [`lavaan` [@R-lavaan]](https://lavaan.ugent.be)[^lavaan] model syntax:

```{r}
model_truth <-
"MACH =~ 0.27*x1 + 0.22*x2 + 0.28*x3 + 0.36*x4 + 0.41*x5 + 0.5*x6 + 0.43*x7 +
0.32*x8 + 0.36*x9
MACH ~ c(0, 0.2)*1
MACH ~~ c(1, 1)*MACH"
```

[^lavaan]: https://lavaan.ugent.be

Data simulation is particularly easy in lavaan syntax using the function `simulateData`.
More complicated simulations (for example, including missing data simulation) are possible with other packages, e.g., simsem [@simsem], faux [@faux], or simstudy [@goldfeldSimstudyIlluminatingResearch2020].
The following code simulates data from the population model `model_truth` with 1000 cases in each group.

```{r, include=FALSE}
set.seed(1234)
```

```{r}
simulated <- lavaan::simulateData(
  model_truth,
  sample.nobs = c(1000, 1000),
  meanstructure = TRUE,
  standardized = TRUE
)
```

We propose the following procedure for analyzing the data and dealing with possible violations of strong measurement invariance. 
After fitting the model with strong measurement invariance [@flake2020], we automatically free equality constraints, first on loadings, and then on intercepts based on the Lagrange Multiplier test.
The model is iteratively refit and improved until at most three loadings, or three intercepts are freed, or the Lagrange Multiplier test is not significant for the constraints on loadings or intercepts.
While reconsidering our modeling assumptions, it is also unlikely that the data are perfectly normally distributed.
First, we have rating scale items and thus our observations are categorical.
Second, @Jones2013 reported means not in the middle of the scale, thus suggesting some skewness to the items.
For simplicity of our illustration, we do not simulate these likely data problems (but doing so would undoubtedly make the simulation more realistic).

If you are satisfied with your assumptions for the data simulation, it helps to write a function that expresses them, like this one:
<!--Hands-on: build simple functions tailored for your analysis-->

```{r generate_data}
```

And to build another function that captures your planned analysis.
Note that this function is again composed of several functions, which you may inspect on `r link("https://github.com/aaronpeikert/repro-tutorial/blob/main/R/simulation_funs.R", "GitHub")`.

```{r planed_analysis}
```

```{r extract_results}
```

With its help we can simulate 1000 cases and fit our analysis.

```{r}
generate_data(1000) %>% 
  planed_analysis() %>% 
  extract_results()
```

Note, that this certainly results in different numbers each time we run it (see ["Random Number Generation"](#prng)).
Of course it is already helpful to simulate a single dataset to test if the analysis works in principle, but repeating the analysis several times and vary parameters lets you verify more properties of your analysis.
First, you can check if the planned analysis recovers the simulated values (`std_estimate` should be close to the simulated mean difference of 0.2).
Second, it enables you to plan an appropriate sample size because you are able to quantify the sampling variability.

To complete a such a  Monte-Carlo simulation study we need a function that repeats the analysis for different conditions (here only for different n):

```{r simulation_study}
```

Such Monte-Carlo simulation can help to decide between several trade-offs.
Historically, one important trade-off is between sample size and power (the probability to detect an effect of given size if it exists).
However, the question can be put more broadly and it is possible to consider various other trade-offs of relevant study design parameters other than sample size, such as number of measurement occasions [e.g., @brandmaier2015lifespan], reliability of measures [e.g., @brandmaier2018precision], or parameters related to the structure of planned-missing data design [@brandmaier2020optimal].
To show that there can also be alternative approaches for sample size planning other than those that are restricted to null hypothesis testing, we choose to visualize the precision of estimation as the density distribution of the parameter of interest (latent mean difference by gender of Machiavellism) instead of statistical power.
Such process still allow us to formulate expectations for the expected precision of the chosen analysis. For example, with a sample size of 1,000 per group, 80% of the simulated standardized effect sizes fall into a range of +/- .1 standardized difference points. This claim is useful and informative irrespective of whether we embark on null hypothesis testing later on or not.
The R code to reproduce the simulation and the figure can be found in the accompanying [GitHub repository](https://github.com/aaronpeikert/repro-tutorial)[^gh-repo]:.

[^gh-repo]: https://github.com/aaronpeikert/repro-tutorial

```{r, include=FALSE}
res <- read_csv(here::here("data", "simulation_results.csv"), col_types = "ddddd")
```

```{r hash}
```

```{r, include=FALSE}
res_interval <- res %>%
  select(n_obs, std_estimate) %>% 
  filter(!is.na(std_estimate)) %>% 
  group_by(n_obs) %>% 
  summarise(interval = list(interval(std_estimate, seq(0.05, .2, 0.01))), 
            .groups = "drop", median = median(std_estimate)) %>% 
  unnest(c(interval)) %>% 
  mutate(width = upper - lower, perc = 1 - alpha) %>% 
  group_by(alpha) %>% 
  arrange(n_obs) %>% 
  mutate(across(c(width, upper, lower),
                ~slide_dbl(.x, mean, na.rm = TRUE, .before = 2, .after = 0, .complete = FALSE))) %>% 
  ungroup() %>% 
  arrange(perc)
```

```{r, include=FALSE}
choosen_perc <- .8
choosen_width <- .2
true_diff <- .2
est_diff <- res %>% filter(n_obs == max(n_obs)) %>% pull(std_estimate) %>% mean()
minn <- filter(res_interval, perc == choosen_perc, width < choosen_width) %>% 
  filter(n_obs == min(n_obs))
sim_n <- pull(res, n_obs)
sim_nmax <- max(sim_n)
sim_n_unique <- unique(sim_n)
sim_nstep <- sim_n_unique[2] - sim_n_unique[1]
```

```{r, include=FALSE}
plot_interval <- res_interval %>%
  ggplot(aes(
    n_obs,
    ymin = lower,
    ymax = upper,
    fill = perc,
    group = alpha
  )) +
  geom_ribbon() +
  geom_line(aes(y = median), color = "white") +
  geom_hline(yintercept = true_diff, linetype = "dotted") +
  scale_fill_viridis_c(
    begin = 1,
    end = 0,
    breaks  = 1 - seq(0.05, .2, 0.05),
    labels = . %>% scales::percent(accuracy = 1)
  ) +
  scale_y_continuous(n.breaks = 10) +
  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
  theme_minimal() +
  labs(x = "N per Group", y = "Standardized Effect Size") +
  theme(legend.title = element_blank()) +
  NULL
```

```{r, include=FALSE}
plot_width <- res_interval %>%
  ggplot(aes(n_obs, width, color = perc, group = perc)) +
  geom_line() +
  geom_point(
    data = minn,
    size = 3,
    color = "black",
    shape = 3
  ) +
  scale_color_viridis_c(
    begin = 1,
    end = 0,
    breaks = 1 - seq(0.05, .2, 0.05),
    labels = . %>% scales::percent(accuracy = 1)
  ) +
  theme_minimal() +
  scale_y_log10(breaks = seq(0, 0.8, .1)) +
  scale_x_continuous(breaks = seq(0, sim_nmax, sim_nstep * 10)) +
  labs(x = "N per Group", y = "Width of quantile\n(Standardized Effect Size)") +
  theme(legend.title = element_blank()) +
  NULL
```

```{r, include=FALSE}
simulation_plot <- plot_interval +
  theme(axis.text.x = element_blank(), axis.title.x = element_blank()) + 
  plot_width +
  plot_layout(nrow = 2, guides = 'collect') +
  plot_annotation(tag_levels = 'A')
```

```{r, simulation-plot, echo=FALSE, fig.cap="*Plot A* depicts the range of effect sizes we observed in the simulation, where the colour encodes the density. The light yellow interval shows where 80% of the estimates land. The dotted line represents the simulated true effect size; the white line shows the median observed effect sizes. *Plot B* represents the width of the interval shown in Plot A on a logarithmic scale. The cross shows the minimum required sample size so that 80% of the estimates land in a range of +/- .1.  "}
simulation_plot
```

### Random Number Generation (optional) {#prng}

Strict reproducibility is at stake with any code that is based on the computer's pseudo random number generator (PRNG) because random numbers will almost always differ with each execution of the code.
To consistently get the same results while preserving the pseudo-random property of the generated numbers, it is good practice to set a random seed that sets the PRNG into a restorable state, in R code: `set.seed(1235)`.
Note however that the consistency of a given seed is not necessarily maintained across different R versions because the default algorithm for generating pseudo random numbers may change (see @Peikert2019 for a demonstration with different R versions).
It is a little more complicated to achieve uncorrelated random numbers in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The standard random number generator (RNG) produces high quality random numbers, but is not safe to use in parallelized code (e.g., when the same code is run with different portions of data in a computer cluster).
The problem is that concurrent random number streams are correlated, because they repeat after some time.
Therefore, if you set a seed for one worker and another seed for another worker, you could end up with the same random numbers but with a phase shift.
@lecuyerGoodParametersImplementations1999 proposed a more computationally expensive RNG, designed to generate many uncorrelated streams.
While this RNG is available as a low level interface in R (with `RNGkind("L'Ecuyer-CMRG")`), we rely on the excellent parallelization framework around the [`future`-package](futureverse.org)[^futureverse] to handle this for us across all workers.

<!--
```{r}
# reproducible but unsafe in parallel code
set.seed(1235)
runif(1)
set.seed(1235)
runif(1)
```
-->
By wrapping code that uses the internal random number generator (here, the command to create a random number from the uniform distribution, `runif(1)`) with the `future` approach, we guarantee that the random numbers will differ across parallel executions of that code, will achieve a high quality of randomness, but still will be deterministic and thus reproducible:

```{r}
# reproducible and safe in parallel code
future::value(future::future(runif(1), seed = 1235))
```

[^futureverse]: https://www.futureverse.org

## Preregistration (optional)

<!--Hands-on: build simple functions tailored for your analysis-->

Preregistration is the specification of a research plan in advance of data analysis.
The goal of preregistration is transparency and a clear-cut divide into exploration and confirmation.
Thus, it serves as a tool that may help to increase the credibility of empirical results.
Without getting into too much detail of the meta-scientific arguments for preregistration, we want to note two advantages [@nosekPreregistrationHardWorthwhile2019].
First, a preregistration is an opportunity to plan a study carefully.
Gathering empirical data requires many resources, and better planning increases the chances that these efforts are worthwhile.
For us, a good preregistration is the same as a good plan; the only difference is that a preregistration is made public beforehand.

>>>>

# Introduction

The papers aim is twofold.
First, we want to introduce the reader to a workflow that ensures reproducibility---defined as the ability to obtain identical results from the *same* data with the *same* computer code. 
The research community has long accepted reproducibility as integral to "good scientific practice."
Often, however, projects can not be reproduced by other research teams, despite the best efforts by the creators.
We present four reasons that lead to irreproducibility, explain four solutions from the software engineering community that eliminate these and show the researchers how the R package `repro` helps them to implement the solutions.
The second objective is to present an alternative preregistration scheme that builds upon a reproducible workflow, the so-called preregistration as code (PAC), where the researchers write the bulk of the manuscript, including the sections Introduction, Methods, and Results---initially based on simulated data, before they gather data.

This tutorial is aimed at researchers who are willing to make code, data, and materials available and should help them to ensure reproducibility of relevant statistical results.
While the willingness to share materials is required for reproducing a project it is not sufficient.
For example, @hardwicke2018 attempted to reproduce results from open materials in the journal Cognition. 
Out of 35 published articles, results of 22 articles could be reproduced but in 11 of these cases, assistance from the original authors was required.
For 13 articles, at least one outcome could not be reproduced—--even with the original authors' assistance.
@obels2020 showed that in 62 Registered Reports, 41 had data available, and 37 had analysis scripts available.
The authors could execute only 31 of the scripts without error and were able to reproduce the results of only 21 articles (within a reasonable period of time).

Even though transparency has increased across scholarly disciplines (that is, data and methods are increasingly openly shared), most of these open repositories do not provide sufficient information to reproduce relevant computational and statistical results.
These failed attempts to reproduce highlight the need for widely accepted reproducibility standards.
To increase the proportion of reproducible projects, such reproducibility standard has to balance the trade-off between being rigorous and being easy to use .
A rigorous standard increases the likelihood of any individual project to be actually reproducible.
An easy to use standard on the other hand is more likely to be employed across research projects.
The remainder of this paper is structured along these lines.
The first part introduces theoretical concepts and software solutions that together form a rigorous reproducibility standard.
The second part is a tutorial that shows how these software solutions can be easily employed by researchers with the help of the R package `repro`.

# Concepts and Software Solutions

<!-- could andreas write this? --->

# Tutorial

# Preregistration as Code

We argue that a preregistration as code (PAC) is an excellent planning tool that offers several benefits over traditional preregistration.
For a PAC, researchers write a reproducible, dynamically generated manuscript including the sections Introduction, Methods, and Results before they gather data.

We consider three criteria when planning a study.
First, the plan should be comprehensive, e.g. state the research question and describe the study design and analysis.
Second, the plan should be effective, meaning that the researchers can reasonably expect it to answer the research question at hand. 
Third, the plan should be efficient; therefore, it only employs the necessary amount of resources.

To ensure a comprehensive plan, we suggest that researchers borrow from widely employed standards for writing an empirical manuscript.
The introduction and theoretical background provide the basis for the studies design and can be written before gathering data.
We argue that researchers can plan more effectively if they formulate the planned analysis as a method section that again follows best practices for describing empirical research.
To ensure that the analysis is technically feasible, they can translate their method section into executable computer code for the data analysis using simulated data.
The researchers can include a "faked " results section based on the simulated data, employing dynamic document generation.
Repeated simulation with varying parameters, a so-called Monte Carlo simulation, can then be employed to plan resources to assure efficiency.
For example, researchers can assess the power (the probability to detect an effect of a given size) for varying numbers of observations in the simulation to determine the required sample size for their study.

Defining the research questions and planning data analysis before observing the research outcomes is called preregistration [https://www.pnas.org/content/115/11/2600].
Preregistration increases the credibility of empirical results in three ways [preregistration is hard and worthwhile, Nosek].

> First, preregistration of analysis plans makes clear which analyses were planned a priori and which were conducted post hoc. This improves calibration of uncertainty for unplanned analyses, and diagnosticity of statistical inferences for planned analyses. [...] Doing so has the benefit of strengthening statistical inferences as compared with unplanned analyses. [...] Second, preregistration enables detection of questionable research practices such as selective outcome reporting (John et al. 2012) or Hypothesizing After the Results are Known (HARKing; Kerr, 1998). Third, preregistration of studies can reduce the impact of publication bias—particularly the prioritization of publishing positive over negative results—by making all planned studies discoverable whether or not they are ultimately published.

A preregistration as code offers four advantages over classical preregistration.
First, a PAC removes any ambiguity regarding the translation of an analysis plan into code.
Second, despite being rigorous, it offers the flexibility to incorporate data-dependent decisions if they can be formulated as code.
Researchers can, for example, formulate conditions under which they prefer one analysis over the other, i.e. if distributional assumptions are not met, employ robust methods or automated variable selection mechanisms.
Third, deviations from the preregistration are more explicit because they are reflected in changes to the code a property facilitated by version control.
Fourth, the preregistration is merely a development stage of the final manuscript, thus saving the researchers from writing and the reviewer from evaluating two separate documents.
